Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-06-17 manuscript_results.Rmd
cleaning code up to line 1600

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
```

@@

```{r}
library(dplyr)
library(ggplot2)

@@

  # Create the box plot with additional facet for categories
plt1<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 5) +
    labs(
      title = "A) Models without log-back transformations",
      x = "",
      y = "Mean Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()+
  theme(legend.position = "none")
plt1
#ggsave(plt1, height = 12, width = 8, filename="fig1A.jpg")  
```

@@

```{r}
i  <- 57
ii <- 8 # state number

@@

# Convert to Date if not already
list_of_states[[ii]]$target_end_date <- as.Date(list_of_states[[ii]]$target_end_date)

@@

```{r}
# 2) build your forecast_df for i=55, ii=8

i  <- 57 # target_end_date
ii <- 8 # state number

@@

conceptual_figure = ggplot() +
  # 3a) two-year observed history as a black line (mapped to color for legend)
  geom_point(data = subset_data, 
            aes(x = target_end_date, y = cases, color = "Fitting data"), 
            size = 2.5) +

  # 3b) forecast ribbon (not in legend)
  geom_ribbon(data = forecast_df, 
              aes(x = Dates, ymin = Lower, ymax = Upper), 
              fill = "gray", alpha = 0.4) +

  # 3c) forecast mean (dashed blue line + hollow points)
  geom_point(data = forecast_df, 
            aes(x = Dates, y = Value, color = "Mean forecast"), 
             size = 3) +

  geom_point(data = forecast_df, 
             aes(x = Dates, y = Value, color = "Mean forecast"), 
             shape = 21, fill = "white", size = 2) +

  # 3d) actual post-cutoff cases (mapped to color for legend)
  geom_point(data = forecast_df, 
             aes(x = Dates, y = Hosp, color = "Hospitalizations"), 
             shape = 16, size = 2.5) +

  # Manually define legend entries
  scale_color_manual(name = "Legend",
                     values = c("Fitting data" = "red", 
                                "Mean forecast" = "cyan3",
                                "Hospitalizations" = "black")) +
    scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +

@@

```
Calculate mean weighted interval score for all forecast weeks on each model.

@@

      ES64_AR_LB = mean(ES64_AR_LB, na.rm = TRUE),
      ES64_ADJ_LB = mean(ES64_ADJ_LB, na.rm = TRUE),
      ES64_EPI_LB = mean(ES64_EPI_LB, na.rm = TRUE),
      ES64_TMP_LB = mean(ES64_TEMP_LB, na.rm = TRUE),
      ES64_AVG_LB = mean(ES64_AVG_LB, na.rm = TRUE),

    )
}

# Now you can use the function with any dataframe
W1_Lg <- calculate_mean_wis(filtered_df_W1)
W2_Lg <- calculate_mean_wis(filtered_df_W2)
W3_Lg <- calculate_mean_wis(filtered_df_W3)
W4_Lg <- calculate_mean_wis(filtered_df_W4)

@@

@@

            W4_Lg$AUTO_AR_LB, W4_Lg$ES27_AR_LB, W4_Lg$ES64_AR_LB, 
            W4_Lg$AUTO_ADJ_LB, W4_Lg$ES27_ADJ_LB, W4_Lg$ES64_ADJ_LB, 
            W4_Lg$AUTO_TMP_LB, W4_Lg$ES27_TMP_LB, W4_Lg$ES64_TMP_LB, 
            W4_Lg$AUTO_EPI_LB, W4_Lg$ES27_EPI_LB, W4_Lg$ES64_EPI_LB,
            W4_Lg$AUTO_AVG_LB, W4_Lg$ES27_AVG_LB, W4_Lg$ES64_AVG_LB)

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[2], "ES64_AR_LB" = colors_ar[3],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[2], "ES64_EPI_LB" = colors_epi[3],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[2], "ES64_TMP_LB" = colors_tmp[3],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[2], "ES64_ADJ_LB" = colors_adj[3],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[2], "ES64_AVG_LB" = colors_avg[3]
)

  # Create the box plot with additional facet for categories
plt2<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 5) +
    labs(
      title = "B) Models with log-back transformations",
      x = "",
      y = "Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()+
  theme(legend.position = "none")
plt2
#ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
```

# Combine in a 2x2 grid
combined_plot <- (plt1 | plt2) + 
  plot_annotation(
    title = "Mean WIS for the 24 models",
    subtitle = "Results by target week",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))
  )
# Print
print(combined_plot)

# Save to file
ggsave("Fig3.jpg", combined_plot, width =9, height =10)
```

COMBINING THE RESULTS WITH LOG AND NO LOG TRASNFORMATION

@@

Here I include a column with the best model result based on the lowest mean(WIS) of each state.

@@

# Extract the columns of interest 
cols <- colnames(W1)[-1]
# Initialize a vector to store results
W1$Best_Result <- character(nrow(W1))

# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W1)) {
  # Find the column name with the minimum value for each row
  W1$Best_Result[i] <- cols[which.min(W1[i, cols])]
}

# REORDER BY FREQUENCY
W1$Best_Result <- fct_infreq(W1$Best_Result)
# Create a new column to classify the model type
W1$Model_Type <- ifelse(grepl("_LB$", W1$Best_Result), "With log-back transformation", "Without log-back transformation")

@@

# Extract the columns of interest 
cols <- colnames(W2)[-1]
# Initialize a vector to store results
W2$Best_Result <- character(nrow(W2))
# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W2)) {
  # Find the column name with the minimum value for each row
  W2$Best_Result[i] <- cols[which.min(W2[i, cols])]
}

# REORDER BY FREQUENCY
W2$Best_Result <- fct_infreq(W2$Best_Result)
# Create a new column to classify the model type
W2$Model_Type <- ifelse(grepl("_LB$", W2$Best_Result),  "With log-back transformation","Without log-back transformation")

@@

######################################
# BEST RESULT
# Extract the columns of interest 
cols <- colnames(W3)[-1]
# Initialize a vector to store results
W3$Best_Result <- character(nrow(W3))

# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W3)) {
  # Find the column name with the minimum value for each row
  W3$Best_Result[i] <- cols[which.min(W3[i, cols])]
}

# REORDER BY FREQUENCY
W3$Best_Result <- fct_infreq(W3$Best_Result)
# Create a new column to classify the model type
W3$Model_Type <- ifelse(grepl("_LB$", W3$Best_Result),  "With log-back transformation","Without log-back transformation")

@@

###################################
# BEST RESULT
# Extract the columns of interest
cols <- colnames(W4)[-1]
# Initialize a vector to store results
W4$Best_Result <- character(nrow(W4))
# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W4)) {
  # Find the column name with the minimum value for each row
  W4$Best_Result[i] <- cols[which.min(W4[i, cols])]
}
W4$Best_Result <- fct_infreq(W4$Best_Result)
# Create a new column to classify the model type
W4$Model_Type <- ifelse(grepl("_LB$", W4$Best_Result),  "With log-back transformation","Without log-back transformation")
# Print merged results
head(W4)

@@

```{r}
custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[1], "ES64_AR_LB" = colors_ar[1],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[1], "ES64_EPI_LB" = colors_epi[1],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[1], "ES64_TMP_LB" = colors_tmp[1],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[1], "ES64_ADJ_LB" = colors_adj[1],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[1], "ES64_AVG_LB" = colors_avg[1],

@@

# define a manual pattern scale
custom_patterns <- c(
  # AR
  "AUTO_AR"      = "circle",      "AUTO_AR_LB"      = "circle",
  "ES27_AR"      = "stripe",      "ES27_AR_LB"      = "stripe",
  "ES64_AR"      = "crosshatch",  "ES64_AR_LB"      = "crosshatch",
  # EPI
  "AUTO_EPI"     = "circle",      "AUTO_EPI_LB"     = "circle",
  "ES27_EPI"     = "stripe",      "ES27_EPI_LB"     = "stripe",
  "ES64_EPI"     = "crosshatch",  "ES64_EPI_LB"     = "crosshatch",
  # TMP
  "AUTO_TMP"     = "circle",      "AUTO_TMP_LB"     = "circle",
  "ES27_TMP"     = "stripe",      "ES27_TMP_LB"     = "stripe",
  "ES64_TMP"     = "crosshatch",  "ES64_TMP_LB"     = "crosshatch",
  # ADJ
  "AUTO_ADJ"     = "circle",      "AUTO_ADJ_LB"     = "circle",
  "ES27_ADJ"     = "stripe",      "ES27_ADJ_LB"     = "stripe",
  "ES64_ADJ"     = "crosshatch",  "ES64_ADJ_LB"     = "crosshatch",
  # AVG
  "AUTO_AVG"     = "circle",      "AUTO_AVG_LB"     = "circle",
  "ES27_AVG"     = "stripe",      "ES27_AVG_LB"     = "stripe",
  "ES64_AVG"     = "crosshatch",  "ES64_AVG_LB"     = "crosshatch"
)

#custom_patterns <- custom_patterns[custom_patterns != "circle"]

plot1 <- ggplot(W1, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,            # no border
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45             # set globally, not inside aes()
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W1)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x") +
  guides(pattern_angle = "none")  # this line is optional now but safe

@@

plot2 <- ggplot(W2, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,            # no border
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45             # set globally, not inside aes()
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W2)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x") +
  guides(pattern_angle = "none")  # this line is optional now but safe

@@

plot3 <- ggplot(W3, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,            # no border
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45             # set globally, not inside aes()
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W3)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x") +
  guides(pattern_angle = "none")  # this line is optional now but safe

plot3

plot4 <- ggplot(W4, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,            # no border
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45             # set globally, not inside aes()
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W4)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x") +
  guides(pattern_angle = "none")  # this line is optional now but safe

plot4
```

@@

# Assign week labels
W1$Week_Ahead <- "W1"
W2$Week_Ahead <- "W2"
W3$Week_Ahead <- "W3"
W4$Week_Ahead <- "W4"

# Combine all weeks
all_weeks <- bind_rows(W1, W2, W3, W4)

>>>>

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales") 
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
library("lme4")
library("ez")
```

@@

@@

# Create the box plot with facet for categories
plt1<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 5) +
    labs(
      title = "A) Models without log-back transformations",
      x = "",
      y = "Mean Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  
    theme_minimal()+
  theme(legend.position = "none")
plt1
#ggsave(plt1, height = 12, width = 8, filename="fig1A.jpg")  
```

@@

```{r}
i  <- 57 # forecasted week
ii <- 8 # state number

@@

@@

```{r}
i  <- 57 # forecast week
ii <- 8 # state number

@@

conceptual_figure = ggplot() +
  # two-year fitting data
  geom_point(data = subset_data, 
            aes(x = target_end_date, y = cases, color = "Fitting data"), 
            size = 2.5) +
  # grey ribbon 
  geom_ribbon(data = forecast_df, 
              aes(x = Dates, ymin = Lower, ymax = Upper), 
              fill = "gray", alpha = 0.4) +
  # mean forecast 
  geom_point(data = forecast_df, 
            aes(x = Dates, y = Value, color = "Mean forecast"), 
             size = 3) +

  geom_point(data = forecast_df, 
             aes(x = Dates, y = Value, color = "Mean forecast"), 
             shape = 21, fill = "white", size = 2) +
  # actual hospitalizations 
  geom_point(data = forecast_df, 
             aes(x = Dates, y = Hosp, color = "Hospitalizations"), 
             shape = 16, size = 2.5) +
  # Manually define legend entries
  scale_color_manual(name = "Legend",
                     values = c("Fitting data" = "red", 
                                "Mean forecast" = "cyan3",
                                "Hospitalizations" = "black")) +
    scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +

@@

```
Evaluating other forecasts

```{r}
plot_forecast <- function(i, ii,
                          list_of_states,
                          AUTO_AVERAGE_WEEK1_list,
                          AUTO_AVERAGE_WEEK2_list,
                          AUTO_AVERAGE_WEEK3_list,
                          AUTO_AVERAGE_WEEK4_list) {

  # Define date range for 2-year history
  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
  start_date <- end_date - (7 * 103)

  # Subset observed data
  subset_data <- subset(list_of_states[[ii]],
                        target_end_date >= start_date & target_end_date <= end_date)

  # Build forecast dataframe
  forecast_df <- data.frame(
    Dates = c(
      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
    ),
    Hosp = c(
      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
    ),
    Value = c(
      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
    ),
    Lower = c(
      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
    ),
    Upper = c(
      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
    )
  )

  # Build the plot
  ggplot() +
    geom_point(data = subset_data,
               aes(x = target_end_date, y = cases, color = "Fitting data"),
               size = 2.5) +
    geom_ribbon(data = forecast_df,
                aes(x = Dates, ymin = Lower, ymax = Upper),
                fill = "gray", alpha = 0.4) +
    geom_point(data = forecast_df,
               aes(x = Dates, y = Value, color = "Mean forecast"),
               size = 3) +
    geom_point(data = forecast_df,
               aes(x = Dates, y = Value, color = "Mean forecast"),
               shape = 21, fill = "white", size = 2) +
    geom_point(data = forecast_df,
               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
               shape = 16, size = 2.5) +
    scale_color_manual(name = "Legend",
                       values = c("Fitting data" = "red",
                                  "Mean forecast" = "cyan3",
                                  "Hospitalizations" = "black")) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
         subtitle = "Gray ribbon represents 95% confidence interval",
         x = "", y = "Hospitalizations") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, face = "bold"),
          plot.subtitle = element_text(size = 14),
          axis.text.x = element_text(angle = 45, hjust = 1))
}

for (ii in 1:48) {
  for (i in 60:70) {
    # Safely check if index exists
    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
      p <- plot_forecast(i, ii,
                         list_of_states,
                         AUTO_AVERAGE_WEEK1_list,
                         AUTO_AVERAGE_WEEK2_list,
                         AUTO_AVERAGE_WEEK3_list,
                         AUTO_AVERAGE_WEEK4_list)
      # Show the plot
      print(p)
    }
  }
}

Calculate mean weighted interval score for influenza seasons on each model and each state.

@@

      ES64_AR_LB = mean(ES64_AR_LB, na.rm = TRUE),
      ES64_ADJ_LB = mean(ES64_ADJ_LB, na.rm = TRUE),
      ES64_EPI_LB = mean(ES64_EPI_LB, na.rm = TRUE),
      ES64_TMP_LB = mean(ES64_TEMP_LB, na.rm = TRUE),
      ES64_AVG_LB = mean(ES64_AVG_LB, na.rm = TRUE),
    )
}

# Now we can use the function with any dataframe
W1_Lg <- calculate_mean_wis(filtered_df_W1)
W2_Lg <- calculate_mean_wis(filtered_df_W2)
W3_Lg <- calculate_mean_wis(filtered_df_W3)
W4_Lg <- calculate_mean_wis(filtered_df_W4)

@@

```{r}
# Combine the data into one dataframe
combined_data <- data.frame(
  week = rep(c("W1", "W2", "W3", "W4"), each = 48 * 15),

@@

            W4_Lg$AUTO_AR_LB, W4_Lg$ES27_AR_LB, W4_Lg$ES64_AR_LB, 
            W4_Lg$AUTO_ADJ_LB, W4_Lg$ES27_ADJ_LB, W4_Lg$ES64_ADJ_LB, 
            W4_Lg$AUTO_TMP_LB, W4_Lg$ES27_TMP_LB, W4_Lg$ES64_TMP_LB, 
            W4_Lg$AUTO_EPI_LB, W4_Lg$ES27_EPI_LB, W4_Lg$ES64_EPI_LB,
            W4_Lg$AUTO_AVG_LB, W4_Lg$ES27_AVG_LB, W4_Lg$ES64_AVG_LB)
)

# Create a new column for model families
combined_data <- combined_data %>%
  mutate(category = case_when(
    grepl("AR", Models) ~ "AR",
    grepl("EPI", Models) ~ "EPI",
    grepl("TMP", Models) ~ "TMP",
    grepl("ADJ", Models) ~ "ADJ",
    grepl("AVG", Models) ~ "AVG",
  ))

# Define color ramps for each model family
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")
colors_avg <- c("orange","orange2", "orange3")

# Use a custom color on each model
custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[2], "ES64_AR_LB" = colors_ar[3],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[2], "ES64_EPI_LB" = colors_epi[3],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[2], "ES64_TMP_LB" = colors_tmp[3],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[2], "ES64_ADJ_LB" = colors_adj[3],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[2], "ES64_AVG_LB" = colors_avg[3]
)

# Create the box plot with additional facet for categories
plt2<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 5) +
    labs(
      title = "B) Models with log-back transformations",
      x = "",
      y = "Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()+
  theme(legend.position = "none")
plt2
#ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
```
Figure 3

```{r}
# Combine plot in a 2x2 grid
figure3 <- (plt1 | plt2) + 
  plot_annotation(
    title = "Mean WIS for the 24 models",
    subtitle = "Results by target week",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))
  )
# Print
print(figure3)

# Save to file
ggsave("Fig3.jpg", figure3, width =9, height =10)
```

COMBINING THE RESULTS WITH LOG-BACK AND NO LOG-BACK TRASNFORMATION

@@

Here I include a column with the best model based on the lowest mean(WIS) of each state.

@@

cols <- colnames(W1)[-1] # get states names
W1$Best_Result <- character(nrow(W1)) # create an empty column for the best models

# Give me the model with lower WIS value 
for (i in 1:nrow(W1)) {
  # Find the model with the minimum value on each row
  # based on the column name
  # save it in the best result
  W1$Best_Result[i] <- cols[which.min(W1[i, cols])]
}

# REORDER BY FREQUENCY
W1$Best_Result <- fct_infreq(W1$Best_Result)
# Create a new column to indicate if model has a log-back
# transformation or not
W1$Model_Type <- ifelse(grepl("_LB$", W1$Best_Result), "With log-back transformation", "Without log-back transformation")

@@

# Extract the columns of interest 
cols <- colnames(W2)[-1]  # get states names
W2$Best_Result <- character(nrow(W2)) # create an empty column for the best models
# Give me the model with lower WIS value 
for (i in 1:nrow(W2)) {
  # Find the model with the minimum value on each row
  # based on the column name
  # save it in the best result
  W2$Best_Result[i] <- cols[which.min(W2[i, cols])]
}

# REORDER BY FREQUENCY
W2$Best_Result <- fct_infreq(W2$Best_Result)
# Create a new column to indicate if model has a log-back
# transformation or not
W2$Model_Type <- ifelse(grepl("_LB$", W2$Best_Result),  "With log-back transformation","Without log-back transformation")

@@

######################################
# BEST RESULT
# Extract the columns of interest 
cols <- colnames(W3)[-1] # get states names
W3$Best_Result <- character(nrow(W3))# create an empty column for the best models
# Give me the model with lower WIS value
for (i in 1:nrow(W3)) {
  # Find the model with the minimum value on each row
  # based on the column name
  # save it in the best result
  W3$Best_Result[i] <- cols[which.min(W3[i, cols])]
}

# REORDER BY FREQUENCY
W3$Best_Result <- fct_infreq(W3$Best_Result)
# Create a new column to indicate if model has a log-back
# transformation or not
W3$Model_Type <- ifelse(grepl("_LB$", W3$Best_Result),  "With log-back transformation","Without log-back transformation")

@@

###################################
# BEST RESULT
# Extract the columns of interest 
cols <- colnames(W4)[-1] # get states names
W4$Best_Result <- character(nrow(W4)) # create an empty column for the best models

# Give me the model with lower WIS value
for (i in 1:nrow(W4)) {
  # Find the model with the minimum value on each row
  # based on the column name
  # save it in the best result
  W4$Best_Result[i] <- cols[which.min(W4[i, cols])]
}
W4$Best_Result <- fct_infreq(W4$Best_Result)
# Create a new column to indicate if model has a log-back
# transformation or not
W4$Model_Type <- ifelse(grepl("_LB$", W4$Best_Result),  "With log-back transformation","Without log-back transformation")
# Print merged results
head(W4)

@@

# define colors for the plot
custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[1], "ES64_AR_LB" = colors_ar[1],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[1], "ES64_EPI_LB" = colors_epi[1],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[1], "ES64_TMP_LB" = colors_tmp[1],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[1], "ES64_ADJ_LB" = colors_adj[1],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[1], "ES64_AVG_LB" = colors_avg[1],

@@

# define patterns for the plot
custom_patterns <- c(
  # AR
  "AUTO_AR"      = "circle",      "AUTO_AR_LB"      = "circle",
  "ES27_AR"      = "stripe",      "ES27_AR_LB"      = "stripe",
  "ES64_AR"      = "crosshatch",  "ES64_AR_LB"      = "crosshatch",
  # EPI
  "AUTO_EPI"     = "circle",      "AUTO_EPI_LB"     = "circle",
  "ES27_EPI"     = "stripe",      "ES27_EPI_LB"     = "stripe",
  "ES64_EPI"     = "crosshatch",  "ES64_EPI_LB"     = "crosshatch",
  # TMP
  "AUTO_TMP"     = "circle",      "AUTO_TMP_LB"     = "circle",
  "ES27_TMP"     = "stripe",      "ES27_TMP_LB"     = "stripe",
  "ES64_TMP"     = "crosshatch",  "ES64_TMP_LB"     = "crosshatch",
  # ADJ
  "AUTO_ADJ"     = "circle",      "AUTO_ADJ_LB"     = "circle",
  "ES27_ADJ"     = "stripe",      "ES27_ADJ_LB"     = "stripe",
  "ES64_ADJ"     = "crosshatch",  "ES64_ADJ_LB"     = "crosshatch",
  # AVG
  "AUTO_AVG"     = "circle",      "AUTO_AVG_LB"     = "circle",
  "ES27_AVG"     = "stripe",      "ES27_AVG_LB"     = "stripe",
  "ES64_AVG"     = "crosshatch",  "ES64_AVG_LB"     = "crosshatch"
)

plot1 <- ggplot(W1, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,         
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45          
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W1)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x")

@@

plot2 <- ggplot(W2, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,         
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45          
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W2)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x")

@@

plot3 <- ggplot(W3, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,           
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45      
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W3)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x") 
plot3

plot4 <- ggplot(W4, aes(x = Best_Result,
                        pattern = Best_Result,
                        fill    = Best_Result)) +
  geom_bar_pattern(
    colour        = NA,       
    pattern_fill  = "black",
    pattern_size  = 0.2,
    pattern_angle = 45          
  ) +
  scale_fill_manual(values   = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model on the 48 U.S. states (W4)",
       x = NULL, y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_x") 
plot4
```

Now I will combine all these plots.

```{r}
# Include columns with target week labels
W1$Week_Ahead <- "W1"
W2$Week_Ahead <- "W2"
W3$Week_Ahead <- "W3"
W4$Week_Ahead <- "W4"

# Combine all dataframes
all_weeks <- bind_rows(W1, W2, W3, W4)

# Create combined plot
combined_plot <- ggplot(all_weeks, aes(x = Best_Result, 
                                       fill = Best_Result,
                                       pattern = Best_Result)) +
  geom_bar_pattern(
    pattern_fill  = "black",
    pattern_size  = 0.1,
    pattern_angle = 45,  
    colour        = NA
  ) +
  scale_fill_manual(values = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +
  labs(title = "Best model based on the lowest mean WIS",
       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
       x = "", y = "Number of states",
       fill = "Best model",
       pattern = "Best model") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold"),legend.position = "left") +
  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x")

```
These legends from ggplot don't work well. I will create an legend for a random plot and I will use this legend in the previous plot.

@@

# Define the custom patterns
custom_patterns <- c("AUTO" = "circle", "ES27" = "stripe", "ES64" = "crosshatch")
custom_fills <- c("AUTO" = "white", "ES27" = "white", "ES64" = "white")

# Create a example of plot
legend_data <- data.frame(
  Model_Type = c("AUTO", "ES27", "ES64", "ES27", "ES64"),
  Model_Family = c("AR","TMP", "EPI", "ADJ","AVG"),
  x = 1:5,
  y = 1
)

# Define colors
custom_fills <- c("AR" = colors_ar[1], "TMP"=colors_tmp[1], "EPI" = colors_epi[1], "ADJ" = colors_adj[1], "AVG" = colors_avg[1])
